---
title: "analysis"
format: pdf
author: "Group Members: Ryan and Veronica"
---
Github repository: [View our GitHub repository](https://github.com/rymooney/stat-learning-group-analysis)

```{r}
hallucination_cases <- read.csv("hallucination_cases.csv")
```

```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
```

The `Professional.Sanction` variable describes if a professional sanction, official penalties or disciplinary measures imposed on a licensed worker for an ethical violation, was given. I am curious if having a misrepresented hallucination in the case correlates with an increase in a professional sanction being issued.
```{r}
analysis_data <- hallucination_cases |>
  mutate(
    misrepresented = str_detect(
      `Hallucination.Items`,
      regex("Misrepresented:", ignore_case = TRUE)
    )
  )

ggplot(analysis_data, aes(x = misrepresented, fill = `Professional.Sanction`)) +
  geom_bar(position = "fill") +
  labs(
    x = "Misrepresented Hallucination",
    y = "Proportion of Cases",
    fill = "Professional Sanction"
  ) +
  scale_x_discrete(labels = c("FALSE" = "No", "TRUE" = "Yes")) +
  theme_minimal()
```
This gives the proportion of cases receiving a professional sanction among cases with and without a misrepresented hallucination.

Since Professional Sanction is a binary response variable, we can perform logistic regression.

```{r}
analysis_data <- analysis_data |>
  mutate(
    sanction = if_else(`Professional.Sanction` == "Yes", 1, 0)
  )

model <- glm(
  sanction ~ misrepresented,
  data = analysis_data,
  family = binomial
)

summary(model)
```

Because logistic regression coefficients are in log-odds, exponentiating the coefficient gives the odds ratio. 
```{r}
exp(0.5532)
```

The logistic regression estimated that cases with a misrepresented hallucination had 1.74 times the odds of receiving a professional sanction compared with cases without one ($\beta_1$ = 0.5532, p = 0.000746). A logistic regression model is appropriate because the response variable, `Professional.Sanction`, is binary, and logistic regression is designed to model the probability of a binary outcome based on the predictor variables. A limitation is that the model only includes whether a case contained a misrepresented hallucination and does not account for other factors  such as the severity or number of hallucinations, the type of  proceeding, or other potentially significant explanatory variables.

